---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

library(utils)
library(LambertW)
library(optband)

```

## README

This package instead approaches simultaneous confidence bands for survival functions purely from an optimization perspective: given a certain coverage level, obtain bands such that the area between is minimized. `optband` imports `LambertW` and `utils` and provides an approximate solution based off local time arguments for both the survival and cumulative-hazard functions using via the `opt.ci` function.

## Installation

````{r, message = FALSE}
install.packages("devtools", repos="http://cran.rstudio.com/")
library(devtools)
devtools::install_github("seasamgo/optband")
library(optband)
````

## Methods 

`opt.ci(survi, conf.level = 0.95, fun = 'surv', tl = NA, tu = NA)`

`opt.ci` takes a `survfit` object from the `survival` package with the desired $1-\alpha$ coverage level, function of interest (either `'surv'` for the survival function or `'cumhaz'` for the cumulative-hazard function), and optional upper or lower bounds for data truncation. Defaults are $\alpha=0.05$, `fun = 'surv'`, `tl = NA`, `tu = NA`.

Other methods/functions are internal and include `cumhaz.var`, `modify.surv.fun`, `psi`, `riemsum`, and `surv.range`. For more, please view their corresponding help file.

## Example

Obtain confidence band for bladder cancer data set inherent to the `survival` package:

````{r}
library(survival)
S <- survival::survfit(Surv(stop, event) ~ 1, type = "kaplan-meier", data = bladder)
opt.S <- optband::opt.ci(S)
````
